Vibrations and Heat Diffusion on the Unit Disc

### 2020-04-18 writeup/draft2.tex
minor edits
more edit

\documentclass{paper}
\usepackage{changepage}
\usepackage{noah}

@@

\author{Noah Jeffrey Blair}
\title{Vibrations and Heat Diffusion on the Unit Disc}
\begin{document}
\maketitle
\tableofcontents
\begin{abstract}

@@

@@

It is known that the integer order Bessel functions have an infinite number of zeros, so we will let $k = \alpha_{nm}$ denote the $m$-th zero of the $n$-th order Bessel function.\\

@@

  (\psi_{n'm'},\psi_{nm}) = \delta_{n'n}\delta_{m'm} \pi J_{n+1}( \alpha_{nm } )

@@

\section{Matching Initial Conditions}
\section{A Particular Solution}

\newpage
\section{References}
Asmar, N. H. (2002). \textit{Applied complex analysis with partial differential equations}. New Jersey: Prentice-Hall Inc.\\
\begin{adjustwidth}{2 cm}{}
This book was used to understand differential equations in polar and spherical coordinate systems, as well as certain properties of Bessel functions.\\
\end{adjustwidth}
Farlow, S. J. (2016). \textit{Partial differential equations for scientists and engineers}. New York: Dover Publications, Inc.\\
\begin{adjustwidth}{2 cm}{}
This book was used to understand the basic types of partial differential equations, as well as some methods for solving them in one dimension. It also introduced the physical interpretation of these equations.\\
\end{adjustwidth}
Fetter, A. L., \& Walecka, J. D. (2003). \textit{Theoretical mechanics of particles and continua}. Mineola, NY: Dover Publications. \\
\begin{adjustwidth}{2 cm}{}
This book was used to understand the physical meaning behind the wave and heat equation, as well as some theorems from Sturm-Liouville theory.\\
\end{adjustwidth}
Tolstov, G. P. (2009). \textit{Fourier series}. New York: Dover Publications, Inc.\\
\begin{adjustwidth}{2 cm}{}
This book was used to understand basic properties of Fourier Series and Fourier-Bessel Series, as well as general methods of eigenfunction expansions.
\end{adjustwidth}
end{document}

>>>>

\documentclass{article}
\usepackage{changepage}
\usepackage{noah}

@@

\author{Noah Jeffrey Blair}
\title{Vibrations and Heat Diffusion on the Unit Disc}
\begin{document}
\maketitle
\begin{abstract}
  The method of eigenfunction expansion is developed and demonstrated for solutions to the Helmholtz equation on the unit disc. Through this method, Greens' functions for the wave and heat equation are generated and particular solutions are analyzed, to show the variety of dynamics possible within the realm of partial differential equations.

@@

\tableofcontents
\newpage

@@

It is known that the integer order Bessel functions have an infinite number of zeros, so we will let $k = \alpha_{nm}$ denote the $m$-th zero of the $n$-th order Bessel function. There is no closed form solution for these zeros, but they can be calculated numerically. In Mathematica, the code is $\alpha_{nm} = \texttt{BesselJZero[n,m]}$.\\

@@

  (\psi_{n'm'},\psi_{nm}) = \delta_{n'n}\delta_{m'm} \pi J_{n+1}( \alpha_{nm } )^2

@@

Our goal now is to use a method similar to the method for Fourier series and Fourier-Bessel series to expand any function into an infinite series of these normal modes. This is known as an eigenfunction-expansion, and is valid for expanding solutions to the equation:
  \mc{L} y_n = \lambda_n y_n
where $\mc{L}$ is an operator that satisfies the following for the given inner product:
  (\mc{L} f,g) = (f,\mc{L} g)
In the case of our normal modes, we hope to write any function $F:B^2\rightarrow\bb{C}$ as:
    F = \sum_{m = 1}^\infty\sum_{n = -\infty}^\infty c_{nm} \psi_{nm} = \sum_{nm} c_{nm}\psi_{nm}
If we take the inner product between one of our basis functions, and solve for $c_{nm}$ we obtain the following equation:
    c_{nm} = \frac{1}{\pi J_{n+1}(\alpha_{nm})^2}\int_0^{2\pi} d\phi\int_0^1 dr F(r,\phi) r e^{- i n \phi}J_n(\alpha_{nm} r)
\section{Matching Initial Conditions}
\section{A Particular Solution}

\newpage
\section{References}
Asmar, N. H. (2002). \textit{Applied complex analysis with partial differential equations}. New Jersey: Prentice-Hall Inc.\\
\begin{adjustwidth}{2 cm}{}
This book was used to understand differential equations in polar and spherical coordinate systems, as well as certain properties of Bessel functions.\\
\end{adjustwidth}
Farlow, S. J. (2016). \textit{Partial differential equations for scientists and engineers}. New York: Dover Publications, Inc.\\
\begin{adjustwidth}{2 cm}{}
This book was used to understand the basic types of partial differential equations, as well as some methods for solving them in one dimension. It also introduced the physical interpretation of these equations.\\
\end{adjustwidth}
Fetter, A. L., \& Walecka, J. D. (2003). \textit{Theoretical mechanics of particles and continua}. Mineola, NY: Dover Publications. \\
\begin{adjustwidth}{2 cm}{}
This book was used to understand the physical meaning behind the wave and heat equation, as well as some theorems from Sturm-Liouville theory.\\
\end{adjustwidth}
Tolstov, G. P. (2009). \textit{Fourier series}. New York: Dover Publications, Inc.\\
\begin{adjustwidth}{2 cm}{}
This book was used to understand basic properties of Fourier Series and Fourier-Bessel Series, as well as general methods of eigenfunction expansions.
\end{adjustwidth}
\end{document}
